---
title: "Regression Models, Course Project"
author: "Aditya Salapaka"
date: "Tuesday 28 April 2015"
output: html_document
---

# Regression Models, Course Project
The objective is to explore the relationship between a set of variables and MPG,
and attempt to answer the following questions:

* “Is an automatic or manual transmission better for MPG?”

* "Quantify the MPG difference between automatic and manual transmissions"

## Executive Summmary
After some basic exploratory analyses, a regression model was chosen which 
quantified the effect of various factors in `mpg`. From statistical and 
regression analysis, the inference was that manual transmission is better for MPG, 
considering the interaction of other variables such as `cyl`, `wt` and `hp`, 
which significantly affect the MPG.

## About The Dataset
The `mtcars` dataset was extracted from the 1975 Motor Trend US magazine, and
comprises fuel consumption and ten aspects of automobile design and performance
for 32 automobiles (1973-74 models).

## Data Transformation

The dataset was loaded into R, and the appropriate variables were factored.

```{r}
data(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am, labels = c("Automatic", "Manual"))
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```

## Exploratory Data Analysis

Following is a pairs plot of the `mtcars` dataset.

```{r cache=TRUE, echo=FALSE}
require("GGally")
ggpairs(mtcars, axisLabels = "internal")
```

The pairs plot shows that many variables have a correlation with mpg. 

Following is a boxplot between `mpg` and `am`.

```{r}
require("ggplot2")
g <- ggplot(mtcars, aes(am, mpg))
g <- g + geom_boxplot() + xlab("Transmission Type") + ylab("MPG") + 
        geom_jitter()
print(g)
```

It is clear that manual transmission corresponds to an increase in mpg. This 
will be proven with a suitable regression model.

## Regression Analysis
Since it is not known which variables have a significant impact, a specific 
model cannot be chosen immediately. For this, a base model correlating all 
variables with `mpg` will be created, and the `step` function will iterate 
multiple regression models to choose the best one.

```{r results='hide'}
initial <- lm(mpg ~ ., data = mtcars)
best <- step(initial, direction = "both")
```

```{r}
summary(best)
```

The best model has `mpg` depending on `cyl`, `hp`, `wt` and `am`. It has an 
adjusted R-squared value of `r summary(best)$adj.r.squared`, which is the highest achievable value. 

To check that variables other than `am` also contribute to a change in `mpg`,
the `anova` test can be run. The null hypothesis is that `cyl`, `hp`, `wt` and
`am` do not affect `mpg`.

```{r}
base <- lm(mpg ~ am, data = mtcars)
anova(base, best)
```

A p-value of 1.69e-08 helps rejects the null hypothesis.

## Residual Plot and Diagnostics
Following is a plot of the best model.
```{r echo=FALSE}
par(mfrow = c(2,2))
plot(best)
```

The residuals vs fitted plot is very scattered, indicative of independence.
The normal Q-Q plot is a straight line, indicative of the normality of the data.

### Outlier Analysis
```{r}
tail(sort(hatvalues(best)), 3)
tail(sort(dfbetas(best)[,6]), 3)
```

The results are the same as seen in the plot, which verifies the analysis.

## Statistical Inference
Assuming normality of the data, t-test is performed, with the null hypothesis 
that automatic or manual transmission have no effect on `mpg`.

```{r}
t.test(mpg ~ am, data = mtcars)
```

A p-value of `r t.test(mpg ~ am, data = mtcars)$p.value` helps in rejecting the
null hypothesis. The difference in `mpg` based on transmission is as follows.

```{r echo=FALSE}
t.test(mpg ~ am, data = mtcars)$estimate
```
## Conclusion
The objective was to answer two questions:

### “Is an automatic or manual transmission better for MPG?”
**Manual Transmission is better for MPG.** The boxplot and t.test show that 
automatic transmission gives an `mpg` of 
`r t.test(mpg ~ am, data = mtcars)$estimate[1]`, while manual transmission gives
an `mpg` of `r t.test(mpg ~ am, data = mtcars)$estimate[2]`. 

### "Quantify the MPG difference between automatic and manual transmissions"

Based on the regression model obtained, it can be concluded that:

* `mpg` increases by `r best$coefficients[6]` in Manual Transmission, when 
considering the effect of `cyl`, `hp` and `wt`.